Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2020-10-27 paper/main.tex
fix intro

In this work, we propose and analyze Unbalanced Mallows Model (UMM), a
population-based probabilistic algorithm specifically designed for optimizing
black-box problems over a permutation landscape under a limited budget. There
exists a wide body of research on probabilistic models for permutations including  Plackett-Luce and Mallows
algorithms~\citep{???},\MANUEL{what can we cite from that body of research?: Ekhine will fix it.}
with the most prominent examples being estimation of distribution algorithms
(EDAs)~\cite{Ceberio2013}.  ACO may also be considered a type of EDA, since
it builds a probability distribution model from which solutions are sampled
with a bias towards the best solutions evaluated so far. The proposed UMM
drastically differs from EDAs since EDAs assume that function evaluations are
cheap and they can do many of them. On the contrary, UMM aims at learning as
much as possible with a limited number of function evaluations. Moreover,
unlike EDAs, UMM does not rely on a fixed-size population of solutions that
evolves over ``generations''. Instead, taking inspiration from Bayesian
optimizers, UMM considers a sample of permutations during the whole execution
that is increased by one new permutation at each iteration. Finally,
the balance between intensification and exploration is dynamically adapted by
controlling the variance of the probabilistic Mallows model.  \manuel{}{
  \begin{itemize}
  \item Who else has used Mallows model for black-box optimization?
  \item Who else has used unbalanced borda for optimization?
  \end{itemize}
}

>>>>

In this work, we propose and analyze Unbalanced Mallows Model (UMM), a
population-based probabilistic algorithm specifically designed for optimizing
black-box problems over a permutation landscape under a limited budget. There
exists a wide body of research on probabilistic models for permutations including  Plackett-Luce and Mallows
algorithms~\citep{???},\MANUEL{what can we cite from that body of research?: Ekhine will fix this.}
with the most prominent examples being estimation of distribution algorithms
(EDAs)~\cite{Ceberio2013}.  ACO may also be considered a type of EDA, since
it builds a probability distribution model from which solutions are sampled
with a bias towards the best solutions evaluated so far. The proposed UMM
drastically differs from EDAs since EDAs assume that function evaluations are
cheap and they can do many of them. On the contrary, UMM aims at learning as
much as possible with a limited number of function evaluations. Moreover,
unlike EDAs, UMM does not rely on a fixed-size population of solutions that
evolves over ``generations''. Instead, taking inspiration from Bayesian
optimizers, UMM considers a sample of permutations during the whole execution
that is increased by one new permutation at each iteration. Finally,
the balance between intensification and exploration is dynamically adapted by
controlling the variance of the probabilistic Mallows model.  \manuel{}{
  \begin{itemize}
  \item Who else has used Mallows model for black-box optimization? Ekhine: Va a mirarlo.
  \item Who else has used unbalanced borda for optimization? Ekhine va a añadir 2 frases.
  \end{itemize}
}
